---
title: "Literature review"
author: "Marie Auger-Méthé"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

# Goals

1. Identify the proportion of recent papers that use habitat selection analysis to inform conservation and management
2. Identify the proportion of recent papers that include characteristics of individual, with a focus on health and fitness proxies. 

# General set up

## Load packages

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(here)
library(ggpubr) # To put plots side-by-side
library(grafify) # For color-blind friendly palette
```

# Search and screening processes

A Web of Science search was conducted using the following search term in the Topic field: (resource selection function OR step\*selection function OR habitat selection analys\*) AND (telemetry OR tag\* OR collar\* OR GPS OR Global Positioning System). We used the Web of Science platform to search all databases available from the library of the University of British Columbia, Canada, on April 28, 2025. These databases (with their Web of Science labels in brackets) were: Biological Abstracts (BIOABS), BIOSIS Citation Index (BCI), BIOSIS Previews (BIOSIS), Current Contents Connect (CCC), Data Citation Index (DRCI), Derwent Innovations Index (DIIDW), FSTA - the food science resource (FSTA), Grants Index (GRANTS), KCI-Korean Journal Database (KJD), MEDLINE (MEDLINE), Policy Citation Index (PCI), Preprint Citation Index (PPRN), ProQuest Dissertation & Theses Citation Index (PQDT), Research Commons (RC), SciELO Citation Index (SCIELO), Web of Science Core Collection (WOS), and Zoological Record (ZOOREC). We restricted the search to documents published between April 29, 2024 and April 28, 2025 using the Publication Date fields. 

We screened all results using the following search criteria: the article must use empirical GPS data (fast-loc GPS data included) of non-human animals for more than two individuals and apply a habitat selection analysis based on a used-available approach  (sensu Matthiopoulos et al., 2023; e.g., RSF, SSF, inhomogeneous Poisson point process) to understand habitat selection. The article must be published (preprints were excluded), must contain sufficient information to assess whether individual characteristics were included in the analysis, and must be written in English. 

Devin A. Lyons (DAL) screened all titles and removed duplicates and patents. DAL and Marie Auger-Méthé (MAM) screened the abstracts. MAM screened the full texts that remained after previous exclusions and extracted information from the articles.

# Visualizing the results

## Loading the literature review data

```{r loading_data}
review <- read_excel(path=file.path(here("Data/0_Review"),
                                  "Literature_review.xlsx"))
head(review)
```

We want to extract only the papers that were kept after screening.

```{r screening}
review <- review %>% 
  filter(TitleExclude == "N"  & AbstractExclude == "N" & FullTextExclude == "N")
head(review)
```

Look at number of papers kept.

```{r number_papers}
n <- nrow(review)
n
```

## Use for conservation and management inference

Look at the potential answers to the question: Is conservation or management inference made based on the results?

```{r conservation}
review %>% select("ManagementConservationInference") %>% unique()
```

Look at proportion of the papers with habitat selection analysis that make conservation or management statements based on the results.

```{r prop_cons}
cons_yes <- review %>% filter(ManagementConservationInference == "Yes") %>% nrow()
cons_yes
cons_yes/n
```

How many of the papers that do not make conservation and management statements based on the results are methodological?

```{r prop_cons_method}
review %>% filter(ManagementConservationInference == "No") %>% 
  select("PrimaryFocus") %>% table()
```

Let's plot these results.

```{r plot_cons}
cons_plot <- ggplot() + 
  geom_bar(data = review, aes(x=as.factor(ManagementConservationInference), 
                              fill = as.factor(PrimaryFocus))) +
  theme_classic() +
  xlab("Conservation/management statement made based on results") + 
  ylab("Number of articles") +
  guides(fill = guide_legend(title="Primary focus")) +
  scale_fill_grafify(palette = "muted") +
  theme(axis.title = element_text(size = rel(1)),
        axis.text = element_text(size = rel(0.9)),
        axis.title.x = element_text(margin = margin(t = 5)),
        axis.title.y = element_text(margin = margin(r = 5)),
        legend.title = element_text(size = rel(1)),
        legend.text = element_text(size = rel(0.9)),
        legend.margin = margin(0, 0, 0, 0),
        panel.border = element_rect(colour = "black", fill=NA, linewidth = 0.5),
        axis.line = element_line(colour = 'black', linewidth = 0.1),
        axis.ticks = element_line(colour = 'black', linewidth = 0.1),
        plot.margin = unit(c(1.5, 0, 0, 1), 'lines'))
cons_plot
```

## Inclusion of individual characteristics

Let's first look at fitness/health proxy.

```{r fitness_char}
review %>% select(IncludeCharacteristicsOfFitnessHealthSpecific) %>% table()
```

Let's change the value of `No, but survival analysis` to something more understandable.

```{r survial}
review <- review %>% 
  mutate(IncludeCharacteristicsOfFitnessHealthSpecific = 
           recode(IncludeCharacteristicsOfFitnessHealthSpecific,
                  "No, but survival analysis" = "No, but include survival analysis"))
review %>% select(IncludeCharacteristicsOfFitnessHealthSpecific) %>% table()
```

```{r fitness_details}
fit_yes <- review %>% filter(IncludeCharacteristicsOfFitnessHealthSpecific == "Yes") %>% 
  select(FitnessHealthCharacteristic, MethodCharacteristic, CommentMethodCharacteristic)
nrow(fit_yes)/n
fit_yes
```


```{r fitness_but_details}
review %>% filter(IncludeCharacteristicsOfFitnessHealth == "Yes, but") %>% 
  select(FitnessHealthCharacteristic, MethodCharacteristic, CommentMethodCharacteristic)
```

Now let's look at the other individual characteristics included in habitat selection analyses.

```{r other_char}
review %>% select(IncludeOtherCharacteristic) %>% table()
```

```{r other_yes}
other_yes <- review %>% filter(IncludeOtherCharacteristic == "Yes") %>% 
  select(OtherCharacteristic, MethodCharacteristic, CommentMethodCharacteristic)
nrow(other_yes)/n
```

```{r other_yes_details}
other_yes %>% select(OtherCharacteristic) %>% table()
```

```{r other_yes_details_method}
other_yes %>% select(MethodCharacteristic) %>% table()
```



```{r plot_characteristics}
char_plot <- ggplot() + 
  geom_bar(data = review, aes(x = "1", 
                              fill = as.factor(IncludeCharacteristicsOfFitnessHealthSpecific))) +
  geom_bar(data = review, aes(x = "2", 
                              fill = as.factor(IncludeOtherCharacteristic))) +
  scale_x_discrete(breaks = c("1","2"),
        labels = c("Fitness/health", "Other characteristic")) +
  theme_classic() +
  xlab("Individual characteristics considered") + 
  ylab("Number of articles") +
  guides(fill = guide_legend(title= "Characteristic included in habitat \nselection analysis")) +
  scale_fill_grafify() +
  theme(axis.title = element_text(size = rel(1)),
        axis.text = element_text(size = rel(0.9)),
        axis.title.x = element_text(margin = margin(t = 5)),
        axis.title.y = element_text(margin = margin(r = 5)),
        legend.title = element_text(size = rel(1)),
        legend.text = element_text(size = rel(0.9)),
        panel.border = element_rect(colour = "black", fill=NA, linewidth = 0.5),
        axis.line = element_line(colour = 'black', linewidth = 0.1),
        axis.ticks = element_line(colour = 'black', linewidth = 0.1),
        plot.margin = unit(c(1.5, 0, 0, 1), 'lines'))
char_plot
```



## Combined plots

```{r Fig_S1_review_res, fig.height=6, fig.width=6, dev = 'png'}
ggarrange(cons_plot, char_plot, 
          labels = c("A", "B"), 
          label.x = rep(0.1, 2), 
          label.y = rep(0.975, 2), 
          font.label = list(size = 12, face="plain"),
          nrow = 2, ncol = 1)
```

# Session info

```{r}
sessionInfo()
```

# Literature cited

Matthiopoulos J., Fieberg J., Aarts G. , Beyer H. L. , Morales J. M. , Haydon D. T. (2015). Establishing the link between habitat selection and animal population dynamics. Ecol Monog 85, 413–436. https://doi.org/10.1890/14-2244.1 
