---
title: "Literature review"
author: "Marie Auger-Méthé"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

# Goals

1. Identify the proportion of recent papers that use habitat selection analysis to inform conservation and management
2. Identify the proportion of recent papers that include characteristics of individual, with a focus on health and fitness proxies. 

# General set up

## Load packages

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(here)
library(ggpubr) # To put plots side-by-side
library(grafify) # For color-blind friendly palette
```


# Visualizing the results

## Loading the literature review data

```{r loading_data}
review <- read_excel(path=file.path(here("Data/0_Review"),
                                  "Literature_review.xlsx"))
head(review)
```

We want to extract only the papers that were kept after screening.

```{r screening}
review <- review %>% 
  filter(TitleExclude == "N"  & AbstractExclude == "N" & FullTextExclude == "N")
head(review)
```

Look at number of papers kept.

```{r number_papers}
n <- nrow(review)
n
```

## Use for conservation and management inference

Look at the potential answers to the question: Is conservation or management inference made based on the results?

```{r conservation}
review %>% select("ManagementConservationInference") %>% unique()
```

Look at proportion of the papers with habitat selection analysis that make conservation or management statements based on the results.

```{r prop_cons}
cons_yes <- review %>% filter(ManagementConservationInference == "Yes") %>% nrow()
cons_yes
cons_yes/n
```

How many of the papers that do not make conservation and management statements based on the results are methodological?

```{r prop_cons_method}
review %>% filter(ManagementConservationInference == "No") %>% 
  select("PrimaryFocus") %>% table()
```

Let's plot these results.

```{r plot_cons}
cons_plot <- ggplot() + 
  geom_bar(data = review, aes(x=as.factor(ManagementConservationInference), 
                              fill = as.factor(PrimaryFocus))) +
  theme_classic() +
  xlab("Conservation/management statement made based on results") + 
  ylab("Number of articles") +
  guides(fill = guide_legend(title="Primary focus")) +
  scale_fill_grafify(palette = "muted") +
  theme(axis.title = element_text(size = rel(1)),
        axis.text = element_text(size = rel(0.9)),
        axis.title.x = element_text(margin = margin(t = 5)),
        axis.title.y = element_text(margin = margin(r = 5)),
        legend.title = element_text(size = rel(1)),
        legend.text = element_text(size = rel(0.9)),
        legend.margin = margin(0, 0, 0, 0),
        panel.border = element_rect(colour = "black", fill=NA, linewidth = 0.5),
        axis.line = element_line(colour = 'black', linewidth = 0.1),
        axis.ticks = element_line(colour = 'black', linewidth = 0.1),
        plot.margin = unit(c(1.5, 0, 0, 1), 'lines'))
cons_plot
```

## Inclusion of individual characteristics

Let's first look at fitness/health proxy.

```{r fitness_char}
review %>% select(IncludeCharacteristicsOfFitnessHealthSpecific) %>% table()
```

Let's change the value of `No, but survival analysis` to something more understandable.

```{r survial}
review <- review %>% 
  mutate(IncludeCharacteristicsOfFitnessHealthSpecific = 
           recode(IncludeCharacteristicsOfFitnessHealthSpecific,
                  "No, but survival analysis" = "No, but include survival analysis"))
review %>% select(IncludeCharacteristicsOfFitnessHealthSpecific) %>% table()
```

```{r fitness_details}
fit_yes <- review %>% filter(IncludeCharacteristicsOfFitnessHealthSpecific == "Yes") %>% 
  select(FitnessHealthCharacteristic, MethodCharacteristic, CommentMethodCharacteristic)
nrow(fit_yes)/n
fit_yes
```


```{r fitness_but_details}
review %>% filter(IncludeCharacteristicsOfFitnessHealth == "Yes, but") %>% 
  select(FitnessHealthCharacteristic, MethodCharacteristic, CommentMethodCharacteristic)
```

Now let's look at the other individual characteristics included in habitat selection analyses.

```{r other_char}
review %>% select(IncludeOtherCharacteristic) %>% table()
```

```{r other_yes}
other_yes <- review %>% filter(IncludeOtherCharacteristic == "Yes") %>% 
  select(OtherCharacteristic, MethodCharacteristic, CommentMethodCharacteristic)
nrow(other_yes)/n
```

```{r other_yes_details}
other_yes %>% select(OtherCharacteristic) %>% table()
```

```{r other_yes_details_method}
other_yes %>% select(MethodCharacteristic) %>% table()
```



```{r plot_characteristics}
char_plot <- ggplot() + 
  geom_bar(data = review, aes(x = "1", 
                              fill = as.factor(IncludeCharacteristicsOfFitnessHealthSpecific))) +
  geom_bar(data = review, aes(x = "2", 
                              fill = as.factor(IncludeOtherCharacteristic))) +
  scale_x_discrete(breaks = c("1","2"),
        labels = c("Fitness/health", "Other characteristic")) +
  theme_classic() +
  xlab("Individual characteristics considered") + 
  ylab("Number of articles") +
  guides(fill = guide_legend(title= "Characteristic included in habitat \nselection analysis")) +
  scale_fill_grafify() +
  theme(axis.title = element_text(size = rel(1)),
        axis.text = element_text(size = rel(0.9)),
        axis.title.x = element_text(margin = margin(t = 5)),
        axis.title.y = element_text(margin = margin(r = 5)),
        legend.title = element_text(size = rel(1)),
        legend.text = element_text(size = rel(0.9)),
        panel.border = element_rect(colour = "black", fill=NA, linewidth = 0.5),
        axis.line = element_line(colour = 'black', linewidth = 0.1),
        axis.ticks = element_line(colour = 'black', linewidth = 0.1),
        plot.margin = unit(c(1.5, 0, 0, 1), 'lines'))
char_plot
```



## Combined plots

```{r Fig_S1_review_res, fig.height=6, fig.width=6, dev = 'png'}
ggarrange(cons_plot, char_plot, 
          labels = c("A", "B"), 
          label.x = rep(0.1, 2), 
          label.y = rep(0.975, 2), 
          font.label = list(size = 12, face="plain"),
          nrow = 2, ncol = 1)
```

# Session info

```{r}
sessionInfo()
```

