---
title: "SSF for narwhals"
author: "Marie Auger-Méthé"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up

## Packages

```{r load.packages, message = FALSE}
library(here) # To simplify data sharing
library(tidyverse)
library(ggpubr) # To put plots side-by-side
library(sf) # to handle spatial object
library(amt) # to prep data for ssf (and fit simpler ssf)
library(glmmTMB)
library(terra)
library(tidyterra)
library(kableExtra)
library(modelsummary)
options(modelsummary_factory_default = 'kableExtra') # New version has tinytable default option
options(modelsummary_factory_latex = 'kableExtra')
options(modelsummary_factory_html = 'kableExtra')
library(tinytable)
library(MuMIn) # for AICc
library(predicts) # For sampling starting locations
library(future.apply) # For parallel lapply
library(stringr) # For selection the good covariate name
library(brms) # von mises for UD
library(RSpectra)
library(Matrix)
```

## Narwhal data

```{r}
narwhals <- read.csv(file=file.path(here("Data/Narwhals/Movement"),"narwhals_clean.csv"))
```

Formatting some of the columns.

```{r}
# ID as factor
narwhals$ID <- as.factor(narwhals$ID)

# Make date time as date time object
narwhals$UTC_datetime <- ymd_hms(narwhals$datetime_UTC, tz="UTC")

narwhals$release_datetime_UTC <- ymd_hms(narwhals$release_datetime_UTC)
```

Check number of GPS locations per deployments and create data table.

```{r data.table}
narwhal.data.table <- narwhals %>% group_by(Individual) %>% 
  summarize("Date of capture" = unique(date(release_datetime_UTC)),
            "N. of days" = as.numeric(round(max(UTC_datetime) - min(UTC_datetime),1)),
            "N. of locations" = length(Individual), 
            "Healed scars" = unique(Healed.Scars)) 
narwhal.data.table
```

```{r save.data.table}
cat(kable(narwhal.data.table, 
          format="latex", 
          caption="Narwhal data summary. We present the date of capture, the number of days and locations used in the analyses, and whether the individual had healed scars. Individual 18001 had a shorter time series because its tag only transmitted for 10 days.",
          booktabs = TRUE, linesep = "", row.names = FALSE),
    file="narwhal_data_summary.tex")
```

Create a sf object for plotting. All of the layers should already projected in EPSG:32617 (WGS 84 / UTM zone 17N) coordinate system.

```{r}
chosen.proj <- 32617
```

```{r}
narwhals.sf <- st_as_sf(narwhals, coords = c("X_UTM", "Y_UTM"),
                        remove = FALSE)
st_crs(narwhals.sf) <- chosen.proj
```

Load the land covariate.

```{r load.land, message = FALSE}
land <- st_read(file.path(here("Data/Narwhals/Covariates/NorthernBaffin_UTM"),
                          "NorthernBaffin_UTM.shp"))
```

```{r}
ggplot() +
  geom_sf(data = narwhals.sf, aes(colour=Healed.Scars), size=0.5) +
  geom_sf(data = land, col="green", fill=NA) +
  coord_sf(xlim = range(narwhals.sf$x), ylim = range(narwhals.sf$y))
```


# Function to prep data for ssf


```{r prep.data.fx}
# tracks <- as.data.frame(narwhals.sf)
# cov <- NULL
# fit.col.name <- "Healed.Scars"
# id.col.name <- "ID"
# t.name = "UTC_datetime"
# x.name = "X_UTM"
# y.name = "Y_UTM"
# mask = list(land)
# n.random.extra = 350
# n.avail = 20
# dt <- 60

prep.data.ssf <- function(tracks, cov, fit.col.name = NULL, id.col.name = "id", 
                          x.name = "X_UTM", y.name = "Y_UTM", t.name = "DatetimeLocal",
                          mask = NULL, dt=15, n.random.extra = 50, n.avail = 30){
  
  
  # Create observed steps
  tracks.id <- tracks %>% nest(.by= c(!! sym(id.col.name), !!! syms(fit.col.name)))
  steps.id <- tracks.id %>% 
    mutate(trk = map(data, function(d) {
      d %>% make_track(.x = !! sym(x.name), .y = !! sym(y.name), .t= !! sym(t.name)) %>%
        track_resample(rate = minutes(dt), tolerance = minutes(1)) %>% steps_by_burst()})) 
  
  
  steps.avail <- steps.id
  steps.avail$rp <- NULL
  
  # We want the parameters used to create the control steps to be from the same distribution
  # We do not want the parameters to be taken each individual seperately
  data <- steps.avail %>% 
    dplyr::select(!! sym(id.col.name), !!! syms(fit.col.name), trk) %>%  
    unnest(cols= c(trk)) 
  gamma_par <-  data$sl_ %>% fit_distr("gamma")
  vonmises_par <- data$ta_ %>% fit_distr("vonmises")
  
  for(i in 1:nrow(steps.avail)){
    steps.avail[[i,"rp"]][[1]] <- steps.id[[i,"trk"]][[1]] %>% 
      random_steps(n_control = n.random.extra, 
                   sl_distr = make_gamma_distr(shape = gamma_par$params$shape,
                                               scale = gamma_par$params$scale,
                                               vcov = gamma_par$vcov),
                   ta_distr = make_vonmises_distr(kappa = vonmises_par$params$kappa,
                                                  vcov = vonmises_par$vcov))
  }
  
  
  for(i in seq_along(cov)){
    steps.avail <- steps.avail %>% 
      mutate(rp = map(rp, function(d){
      extract_covariates(d,cov[[i]])}))
  }
  
  steps.avail <- steps.avail %>% 
    dplyr::select(!! sym(id.col.name), !!! syms(fit.col.name), rp) %>%  
    unnest(cols= c(rp))  
  
  # Important: the `step_id_` will have erroneously similar names across individuals. We will change the name here.
  steps.avail$step_id_ <- paste(steps.avail[[id.col.name]], steps.avail$step_id_, sep="_")
  steps.avail.sf <- st_as_sf(filter(steps.avail, !is.na(ta_)), coords = c("x2_", "y2_"), remove=FALSE)
  st_crs(steps.avail.sf) <- chosen.proj
  
  
  # Remove the control points that are in habitat that are masked (e.g., land for narwhals)

  for(i in seq_along(mask)){
    steps.avail.sf$mask <- rowSums(st_intersects(steps.avail.sf, mask[[i]], sparse=FALSE))
    steps.avail.sf <- subset(steps.avail.sf, !mask)
  }
    
   
  steps.avail.11 <- as.data.frame(steps.avail.sf) %>%
      group_by(step_id_) %>% filter(row_number() >= 1 & row_number() <= n.avail + 1)
  
  # Get the log sl and cos ta. And numeric case for poisson distribution
  steps.avail.11$ln_sl <- log(steps.avail.11$sl_)
  steps.avail.11$cos_ta <- cos(steps.avail.11$ta_)
  steps.avail.11$case_num <- as.numeric(steps.avail.11$case_)
  
  return(list(steps.id = steps.id, steps.avail = steps.avail.11, gamma_par=gamma_par, vonmises_par=vonmises_par))
}
```

We are using the land mask and 60 min time interval.

```{r, cache=TRUE}
set.seed(78369)
narwhals.steps.avail <- prep.data.ssf(as.data.frame(narwhals.sf), cov = NULL, 
                                    fit.col.name = "Healed.Scars", 
                                    id.col.name ="ID", t.name = "UTC_datetime", 
                                    mask= list(land), n.random.extra = 1200, n.avail = 50,
                                    dt = 60)
head(narwhals.steps.avail$steps.avail)
```

Should be n.avail. + 1. Needed lots of `r.random.extra` since tiny fjords.

```{r}
nrow(narwhals.steps.avail$steps.avail)/sum(narwhals.steps.avail$steps.avail$case_num == 1) 
```

Let's calculate distance from shore. Distance from shore for habitat selection should be at the end of the step (select where to go), but for the movement kernel it should be at beginning of the step (movement characteristics based on where they are).

```{r, cache=TRUE}
narwhals.avail.sf <- st_as_sf(narwhals.steps.avail$steps.avail, coords = c("x2_", "y2_"), remove=FALSE, crs = chosen.proj)
narwhals.avail.sf.start <- st_as_sf(narwhals.steps.avail$steps.avail, coords = c("x1_", "y1_"), remove=FALSE, crs = chosen.proj)

land.1 <- st_cast(st_union(land),  to = 'MULTILINESTRING') # Need to make it a line to 
narwhals.avail.sf$distShore.end <- as.numeric(st_distance(narwhals.avail.sf, land.1))/1000
narwhals.avail.sf$distShore.start <- as.numeric(st_distance(narwhals.avail.sf.start, land.1))/1000
```

Let's plot distance from shore.

```{r}
ggplot() +
  geom_sf(data = narwhals.avail.sf,
          aes(color = distShore.end), size = 0.5) +
  geom_sf(data = land, col = "green", fill=NA) +
  coord_sf(xlim = range(narwhals.avail.sf$x1_), 
           ylim = range(narwhals.avail.sf$y1_))

ggplot() +
  geom_sf(data = narwhals.avail.sf,
          aes(color = distShore.start), size = 0.5) +
  geom_sf(data = land, col = "green", fill=NA) +
  coord_sf(xlim = range(narwhals.avail.sf$x1_), 
           ylim = range(narwhals.avail.sf$y1_))
```

Format data and binary variable as 0 and 1.

```{r}
all.steps <- as.data.frame(narwhals.avail.sf)
all.steps$Healed.Scars <- as.numeric(all.steps$Healed.Scars)
```

# Fit models

Function to fit an SSF.

```{r}
fit.poisson.rs <- function(formula, data, 
                           start = list(theta = c(log(1e6),0)), 
                           map = list(theta = factor(c(NA, 1)))){
  ssf <- glmmTMB(formula, family=poisson, 
                 data = data, start = start, map = map)
  return(ssf)
}
```


AICc table function

```{r aicc.table.fx}
AICcTable <- function(m, ...){
  namesObj <- names(tibble::lst(m, ...))
  namesObj[1] <- deparse(substitute(m))
  AICc.table <- AICc(m,...)
  row.names(AICc.table) <- namesObj
  AICc.table <- AICc.table %>% mutate(deltaAICc = AICc - min(AICc, na.rm=TRUE))
  AICc.table[order(AICc.table$deltaAICc),]
}
```

Base info for tables of results.

```{r}
title <- "Parameter estimates from the four SSF models applied to the narwhal movement data. The parameter estimates, along with their standard errors, are presented. A * indicates that the p-value comparing this parameter value to 0 is significant at the significance level of 0.05."

coef_map <- c("distShore.end" = "$\\beta_1: d_{n,t,j}$",
              
              #SL
              "sl_" = "$\\beta_{l}: l_{n,t,j}$",
              
              # SL X distshore
              "distShore.start:sl_" = "$\\beta_2: d_{n,t-1,j} l_{n,t,j}$",
              "sl_:distShore.start" = "$\\beta_2: d_{n,t-1,j} l_{n,t,j}$",
              
              #ln(SL)
              "ln_sl" = "$\\beta_{\\ln}: \\ln(l_{n,t,j})$",
              
              # ln(SL) X distshore
              "distShore.start:ln_sl" = "$\\beta_3: d_{n,t-1,j} \\ln(l_{n,t,j})$",
              "ln_sl:distShore.start" = "$\\beta_3: d_{n,t-1,j} \\ln(l_{n,t,j})$",
              
              # ln(SL) X distshore^2
              "I(distShore.start^2):ln_sl" = "$\\beta_4: d_{n,t-1,j}^2 \\ln(l_{n,t,j})$",
              "ln_sl:I(distShore.start^2)" = "$\\beta_4: d_{n,t-1,j}^2 \\ln(l_{n,t,j})$",
              
              # Turning angle
              "cos_ta" = "$\\beta_{\\theta}: \\cos(\\theta_{n,t,j})$",
          
              ## Healed scars
              # distance to shore X healed scars
              "distShore.end:Healed.Scars" = "$\\beta_5: s_n d_{n,t,j}$",
              "Healed.Scars:distShore.end" = "$\\beta_5: s_n d_{n,t,j}$",
              
              # SL X distshore X healed scars
              "distShore.start:sl_:Healed.Scars" = "$\\beta_{6}: s_n d_{n,t-1,j} l_{n,t,j}$",
              "distShore.start:Healed.Scars:sl_" = "$\\beta_{6}: s_n d_{n,t-1,j} l_{n,t,j}$",
              "sl_:Healed.Scars:distShore.start" = "$\\beta_{6}: s_n d_{n,t-1,j} l_{n,t,j}$",
              "sl_:distShore.start:Healed.Scars" = "$\\beta_{6}: s_n d_{n,t-1,j} l_{n,t,j}$",
              "Healed.Scars:sl_:distShore.start" = "$\\beta_{6}: s_n d_{n,t-1,j} l_{n,t,j}$",
              "Healed.Scars:distShore.start:sl_" = "$\\beta_{6}: s_n d_{n,t-1,j} l_{n,t,j}$",
                  
              # Random effects 
              # "SD($x_{dist.e} \\; |$ ID)",
              "SD (distShore.end ID)" = "$\\sigma_{1}: \\beta_{1,n}$"
              )


estimates <- "{estimate}{stars}"
statistics <- "({std.error})"
```

We will look at four SSFs that compare the environmental covariates at locations associated with the observed animal movement at time $t$ (encoded as $y_{n,t,j} = 1$ for $j = 0$) to locations assumed to be available to the animal at that time (encoded as $y_{n,t,j} = 0$ for $j = 1, ..., J$). Here, these SSFs model the locations of $n = 1, ..., 14$ narwhals for $j = 0, 1,..., 50$ used and control locations. The environmental covariate of interest is distance to shore, and we assumed that distance to shore at the end of step $j$ $d_{n,t,j}$ affects the selection function, while distance to shore at the start of a step $j$ $d_{n,t-1,j}$ affects the movement kernel. We use the mixed Poisson ($\text{Po}$) model framework outlined in Muff et al. (2020).

Model 1 (baseline model): Scars ignored

The SSF fitted models the step of individual $n$ at time $t$, $y_{n,t,j}$, which take the value of 1 for used step ($j = 0$) and 0 for control points ($j = 1, ..., J$), as:

$$\mu_{n,t,j} = \exp(\beta_{0,n,t} + \beta_{1,n} d_{n,t,j}  + \\ \beta_{l} l_{n,t,j} + \beta_{2} d_{n,t-1,j} l_{n,t,j} + \\
\beta_{\ln} \ln(l_{n,t,j}) + \beta_3 d_{n,t-1,j} \ln(l_{n,t,j}) + \beta_4 d_{n,t-1,j}^2 \ln(l_{n,t,j}) + \\
\beta_{\theta} \cos(\theta_{n,t,j})), \qquad y_{n,t,j} \sim \text{Po}(\mu_{n,t,j}),$$

where for the $n^{th}$ individual at time $t$ for step $j$, $d_{n,t,j}$ represents the distance to shore, $l_{n,t,j}$ the step length, and  $\theta_{n,t,j}$ the turning angle. The coefficient $\beta_{0,n,t}$ is the step-specific intercept of animal $n$ at time $t$, which is modeled as a random effect $\beta_{0,n,t} \sim \text{N}(0,1000000^2)$, while the coefficients $\beta_{1,n}$ is the coefficient associated with the habitat selection function and is modelled with a random effect for which we estimate the mean and standard deviation: $\beta_{1,n} \sim \text{N}(\beta_1, \sigma_1^2)$. The rest of the terms are associated with the movement kernel, where $\beta_2$, $\beta_{l}$, $\beta_{\ln}$, $\beta_{\theta}$, $\beta_{3}$, and $\beta_4$ are coefficients associated with the step lengths and turning angles. As explained in Avgar et al. (2015), the inclusion of the step length and its natural logarithm is a way to model the step length distribution as a Gamma distribution. Similarly, the inclusion of the cosine of the turning angle is a method to model the turning angle distribution as a Von Mises. Here, $\beta_2$, $\beta_3$, $\beta_4$ allow for the movement kernel, specifically the step length distribution, to vary as a function of distance to shore.

```{r, message = FALSE}
ssf.noscars <- fit.poisson.rs(formula = case_num ~ -1 + 
                                # Habitat selection function
                                distShore.end +
                                # Movement kernel
                                sl_ + distShore.start:sl_ + 
                                ln_sl + distShore.start:ln_sl + I(distShore.start^2):ln_sl +
                                cos_ta + 
                                (1 | step_id_) +
                                (0 + distShore.end | ID), 
                              data = all.steps)
summary(ssf.noscars)
```

There are many warnings (not shown in compiled version), but these are common when fitting a complex glmm. The fact that we get standard errors is a good sign. Let's look if there are convergence issues.

```{r}
ssf.noscars$fit$convergence
```

No convergence problem.

The three other models use healed scars as a health proxy by including interactions in the model with the variable $s_n$, which takes the value of 1 when healed scars are present on the individual $n$ and 0 when individual $n$ has no healed scars.

Model 2: Scar information in the selection function

As a base, we are using model 1 described above. However, to allow for the selection function to differ between individuals with and without scars, we include an interaction between presence of scars, $s_n$, and distance to shore at the end of the step, $d_{n,t,j}$: 

$$\mu_{n,t,j} = \exp(\beta_{0,n,t} + \beta_{1,n} d_{n,t,j} +  \\ \beta_5 s_n d_{n,t,j} + \\ \beta_{l} l_{n,t,j} + \beta_{2} d_{n,t-1,j} l_{n,t,j} + \\
\beta_{\ln} \ln(l_{n,t,j}) + \beta_3 d_{n,t-1,j} \ln(l_{n,t,j}) + \beta_4 d_{n,t-1,j}^2 \ln(l_{n,t,j}) + \\
\beta_{\theta} \cos(\theta_{n,t,j})), \qquad y_{n,t,j} \sim \text{Po}(\mu_{n,t,j}),$$

where $\beta_5$ represents the coefficient quantifying the relationship between the presence of scars and distance to shore.

```{r, message = FALSE}
ssf.sel.healed <- fit.poisson.rs(formula = case_num ~ -1 + 
                                   # Habitat selection
                                   distShore.end + 
                                   distShore.end:Healed.Scars +
                                   # Movement kernel
                                   sl_ + distShore.start:sl_ + 
                                   ln_sl  + distShore.start:ln_sl  + I(distShore.start^2):ln_sl  + 
                                   cos_ta + 
                                   (1 | step_id_) +
                                   (0 + distShore.end | ID), 
                                  data = all.steps)
summary(ssf.sel.healed)
```

There are many warnings (not shown in compiled version), but these are common when fitting a complex glmm. The fact that we get standard errors is a good sign. Let's look if there are convergence issues.

```{r}
ssf.sel.healed$fit$convergence
```

No convergence problem.

Model 3: Scar information in the movement kernel

As a base, we are using model 1 described above. However, to allow for the relationship with speed and distance to shore to to differ between individuals with and without scars, we include an interaction between the presence of scars, $s_n$, step length, $l_{n,t,j}$, and distance to shore at the start of the step, $d_{n,t-1,j}$: 

$$\mu_{n,t,j} = \exp(\beta_{0,n,t} + \beta_{1,n} d_{n,t,j} +  \\ \beta_{l} l_{n,t,j} + \beta_{2} d_{n,t-1,j} l_{n,t,j} + \\ \beta_6 s_n d_{n,t-1,j} l_{n,t,j} + \\
\beta_{\ln} \ln(l_{n,t,j}) + \beta_3 d_{n,t-1,j} \ln(l_{n,t,j}) +  \beta_4 d_{n,t-1,j}^2 \ln(l_{n,t,j}) + \\
\beta_{\theta} \cos(\theta_{n,t,j})), \qquad y_{n,t,j} \sim \text{Po}(\mu_{n,t,j}),$$

where $\beta_6$ quantifies the relationship between the scars, step length, and distance to shore.

```{r, message = FALSE}
ssf.mov.healed <- fit.poisson.rs(formula = case_num ~ -1 + 
                                   # Habitat selection
                                   distShore.end + 
                                   # Movement kernel
                                   sl_ + distShore.start:sl_ + 
                                   distShore.start:Healed.Scars:sl_ + 
                                   ln_sl + distShore.start:ln_sl + I(distShore.start^2):ln_sl + 
                                   cos_ta + 
                                   (1|step_id_) +
                                   (0 + distShore.end | ID), 
                                  data = all.steps)
summary(ssf.mov.healed)
```

There are many warnings (not shown in compiled version), but these are common when fitting a complex glmm. The fact that we get standard errors is a good sign. Let's look if there are convergence issues.

```{r}
ssf.mov.healed$fit$convergence
```

No convergence problem.

Model 4: Scar information in both the selection function and the movement kernel

To allow the presence of scars to affect both the selection function and the movement kernel, we simply combine models 2 and 3:

$$\mu_{n,t,j} = \exp(\beta_{0,n,t} + \beta_{1,n} d_{n,t,j} +  \\ \beta_5 s_n d_{n,t,j} + \\ \beta_{l} l_{n,t,j} + \beta_{2} d_{n,t-1,j} l_{n,t,j} + \\ \beta_6 s_n d_{n,t-1,j} l_{n,t,j} + \\
\beta_{\ln} \ln(l_{n,t,j}) + \beta_3 d_{n,t-1,j} \ln(l_{n,t,j}) +  \beta_4 d_{n,t-1,j}^2 \ln(l_{n,t,j}) + \\
\beta_{\theta} \cos(\theta_{n,t,j})), \qquad y_{n,t,j} \sim \text{Po}(\mu_{n,t,j}).$$

```{r, message = FALSE}
ssf.mov.sel.healed <- fit.poisson.rs(formula = case_num ~ -1 + 
                                       # Habitat selection
                                       distShore.end + distShore.end:Healed.Scars +
                                       # Movement kernel
                                       sl_ + distShore.start:sl_ + 
                                       distShore.start:Healed.Scars:sl_ + 
                                       ln_sl  + distShore.start:ln_sl  + I(distShore.start^2):ln_sl  + 
                                       cos_ta + 
                                       (1|step_id_) +
                                       (0 + distShore.end | ID),
                                     data = all.steps)
summary(ssf.mov.sel.healed)
```

There are many warnings (not shown in compiled version), but these are common when fitting a complex glmm. The fact that we get standard errors is a good sign. Let's look if there are convergence issues.

```{r}
ssf.mov.sel.healed$fit$convergence
```


```{r}
AICcTable(ssf.noscars,
          ssf.sel.healed,
          ssf.mov.healed,
          ssf.mov.sel.healed)
```

The best model is with healed scar only in the movement kernel. The next best model is more complicated and has a difference in AICc of less than 2, so we don't consider it further. The difference in AICc with the baseline model is much greater than 2.

```{r}
models <- list("Ignored" = ssf.noscars,
               "In selection function" = ssf.sel.healed,
               "In movement kernel" = ssf.mov.healed,
               "In both" = ssf.mov.sel.healed)

modelsummary(models, 
             estimate = estimates, 
             statistic = statistics,
             coef_map = coef_map,
             gof_map = NA,
             fmt = fmt_significant(3),
             stars = c("*"=0.05),
             title = title) %>%
  add_header_above(c("Par: Variables" = 1, "Healed scars" = 4))
```

```{r par.est.latex}
# To output for latex
options("modelsummary_format_numeric_latex" = "plain")
table.latex <- modelsummary(models,
             estimate = estimates,
             statistic = statistics,
             coef_map = coef_map,
             gof_map = NA,
             fmt = fmt_significant(3),
             stars = c("*"=0.05),
             title = title,
             output = "narwhal_par_est.tex",
             escape = FALSE) %>%
  add_header_above(c("Par: Variables" = 1, "Healed scars" = 4),
             escape = FALSE) 
cat(table.latex, file="narwhal_par_est.tex")
```

Does having scars in the movement kernel affect the selection relationship with distance to shore?

```{r}
# Without scar info in movement kernel
round(summary(ssf.noscars)$coef$cond["distShore.end", c("Estimate", "Std. Error")],3)
# With scar info in movement kernel
round(summary(ssf.mov.healed)$coef$cond["distShore.end", c("Estimate", "Std. Error")], 3)
```
A bit, but not massively.

# Interpreting the results

Parameters for the gamma distribution used to sample the control points.

```{r}
b1_shape <- narwhals.steps.avail$gamma_par$params$shape
b2_scale <- narwhals.steps.avail$gamma_par$params$scale
b3_kappa <- narwhals.steps.avail$vonmises_par$params$kappa
```

Functions to calculate the estimated parameters of the gamma.

```{r}
# model <- ssf.noscars
# model <- ssf.mov.healed
# dist <- seq(min(all.steps$distShore.start), ceiling(max(all.steps$distShore.start)), by=0.1)
shape.ssf <- function(b1_shape, model, dist){
  gamma.shape.ssf <- b1_shape + 
    summary(model)$coef$cond["ln_sl",1] +
    dist*summary(model)$coef$cond["distShore.start:ln_sl",1]  + 
    #dist*summary(model)$coef$cond["ln_sl:distShore.start",1]  + 
    (dist^2)*summary(model)$coef$cond["ln_sl:I(distShore.start^2)",1] 
  return(gamma.shape.ssf)
}

scale.ssf <- function(b2_scale, model, dist, healed = FALSE){
  sl.healed.dist <- which(str_detect(row.names(summary(model)$coef$cond), "sl_") & 
             str_detect(row.names(summary(model)$coef$cond), "Healed.Scars") & 
             str_detect(row.names(summary(model)$coef$cond), "distShore.start"))
  sl.dist <- which(str_detect(row.names(summary(model)$coef$cond), "sl_") & 
             !str_detect(row.names(summary(model)$coef$cond), "Healed.Scars") & 
             str_detect(row.names(summary(model)$coef$cond), "distShore.start"))
  gamma.scale.ssf <- 1/((b2_scale^(-1)) - 
                          summary(model)$coef$cond["sl_",1] -
                          dist*summary(model)$coef$cond[sl.dist, 1] -
                          dist*ifelse(healed, 
                                      summary(model)$coef$cond[sl.healed.dist,1],
                                      0))
  return(gamma.scale.ssf)
}


# no need for dist, no other covariate has an effect on tangle
kappa.ssf <- function(b3_kappa, model){
  vmises.kappa.ssf <- b3_kappa + summary(model)$coef$cond["cos_ta",1]
  return(vmises.kappa.ssf)
}


# Function for mean step length
sl.ssf <- function(b1_shape, b2_scale, model, dist, healed = FALSE){
  gamma.shape.ssf <- shape.ssf(b1_shape, model, dist)
  gamma.scale.ssf <- scale.ssf(b2_scale, model, dist, healed)
  mean.sl <- gamma.shape.ssf*gamma.scale.ssf
  sd.sl <- sqrt(gamma.shape.ssf*(gamma.scale.ssf^2))
  return(cbind(mean.sl, sd.sl))
}

```


```{r}
pred.col <- c("hotpink", "blue")
names(pred.col) <- c("TRUE", "FALSE")
```


```{r}
dist.seq <- seq(min(all.steps$distShore.start), ceiling(max(all.steps$distShore.start)), by=0.1)

pred.mean.noscars <- sl.ssf(b1_shape, b2_scale, ssf.noscars, dist.seq)
pred.mean.sel.healed <- sl.ssf(b1_shape, b2_scale, ssf.sel.healed, dist.seq)
pred.mean.mov.healed.T <- sl.ssf(b1_shape, b2_scale, ssf.mov.healed, dist.seq, healed = TRUE)
pred.mean.mov.healed.F <- sl.ssf(b1_shape, b2_scale, ssf.mov.healed, dist.seq, healed = FALSE)
pred.mean.mov.sel.healed.T <- sl.ssf(b1_shape, b2_scale, ssf.mov.sel.healed, dist.seq, healed = TRUE)
pred.mean.mov.sel.healed.F <- sl.ssf(b1_shape, b2_scale, ssf.mov.sel.healed, dist.seq, healed = FALSE)

models.pred <- data.frame(cbind(dist.seq, 
                                pred.mean.noscars, 
                                pred.mean.sel.healed,
                                pred.mean.mov.healed.F, 
                                pred.mean.mov.healed.T,
                                pred.mean.mov.sel.healed.F, 
                                pred.mean.mov.sel.healed.T))
colnames(models.pred) <- c("dist.seq", 
                           "noscars.mean", "noscars.sd",
                           "sel.healed.mean", "sel.healed.sd",
                           "mov.healed.F.mean", "mov.healed.F.sd",
                           "mov.healed.T.mean", "mov.healed.T.sd",
                           "mov.sel.healed.F.mean", "mov.sel.healed.F.sd",
                           "mov.sel.healed.T.mean", "mov.sel.healed.T.sd")

# Make predicted step length in km
models.pred[-1] <- models.pred[-1]/1000
```

```{r}
all.steps$scars <- as.logical(all.steps$Healed.Scars)
all.steps$bin <- as.factor(round(all.steps$distShore.start))

dist.max <- max(all.steps$distShore.start)
bb <- 1
breaks <- seq(0, dist.max+bb, bb)

all.steps <- all.steps %>% mutate(new_bin = cut(distShore.start, breaks=breaks),
                                  new_bin_mid = breaks[findInterval(distShore.start, breaks)] + bb/2)


mean.obs.sl.all <- aggregate(sl_ ~ new_bin_mid, subset(all.steps, case_num ==1), mean)
mean.obs.sl.healed.T <- aggregate(sl_ ~ new_bin_mid, subset(all.steps, case_num ==1 & Healed.Scars == 1), mean)
mean.obs.sl.healed.F <- aggregate(sl_ ~ new_bin_mid, subset(all.steps, case_num ==1 & Healed.Scars == 0), mean)

col.all <- "darkgrey"
```

Plot settings.

```{r}
legend.text.size <- 0.8
plot.mar.l <- c(0.5,0.2,0,0.2)
```

Mean step length plots.

```{r}
nudge.v <- 0.15
out.size <- 0.1

mov.plot <- function(names, healed){
  if(!healed){
    m.plot <- ggplot() + 
      geom_boxplot(data = subset(all.steps, case_num ==1), 
                   aes(x=new_bin_mid, y=sl_/1000, group=as.factor(new_bin_mid)), 
                   fill = alpha(col.all,0.5),
                   outlier.size = out.size,
                   linewidth = 0.2) +
      geom_point(data= mean.obs.sl.all, aes(x=new_bin_mid, y = sl_/1000), 
                 col=col.all, pch=17, size = 0.5) +
      geom_line(data= models.pred, aes(x=dist.seq, y = !! sym(paste(names, "mean", sep="."))))
  }else{
    m.plot <- ggplot() + 
      geom_boxplot(data = subset(all.steps, case_num ==1 & Healed.Scars == TRUE), 
                   aes(x=new_bin_mid, y=sl_/1000, group=as.factor(new_bin_mid)), 
                   fill = alpha(pred.col["TRUE"],0.5), 
                   outlier.colour = pred.col["TRUE"], 
                   outlier.size = out.size,
                   position = position_nudge(x=-nudge.v), 
                   width = 0.25,
                   linewidth = 0.2) +
      geom_point(data= mean.obs.sl.healed.T, aes(x=new_bin_mid-nudge.v, y = sl_/1000), 
                 col=pred.col["TRUE"], pch=17, size = 0.5) +
      geom_line(data= models.pred, aes(x=dist.seq, y = !! sym(paste(names, "T.mean", sep="."))), col=pred.col["TRUE"]) +
      geom_boxplot(data = subset(all.steps, case_num ==1 & Healed.Scars == FALSE), 
               aes(x=new_bin_mid, y=sl_/1000, group=as.factor(new_bin_mid)), 
               fill = alpha(pred.col["FALSE"],0.5), 
               outlier.size = out.size,
               outlier.colour = pred.col["FALSE"],
               position = position_nudge(x=+nudge.v), 
               width = 0.25,
               linewidth = 0.2) +
      geom_point(data= mean.obs.sl.healed.F, aes(x=new_bin_mid+nudge.v, y = sl_/1000), 
                 col=pred.col["FALSE"], pch=17, size = 0.5) +
      geom_line(data= models.pred, aes(x=dist.seq, y = !! sym(paste(names, "F.mean", sep="."))), col=pred.col["FALSE"])
  }
  
  m.plot <- m.plot + xlab("Distance to shore (km)") + ylab("Step length (km)") + 
    guides(Healed.T = guide_custom(grid::circleGrob(r = unit(0.1, "cm"),
                                                  gp = grid::gpar(fill = pred.col["TRUE"], col=NA)), 
                                 title = "With healed scars", order = 3, 
                                 theme = theme(legend.title = element_text(size = rel(legend.text.size)),
                                               legend.position = "top")),
           Healed.F = guide_custom(grid::circleGrob(r = unit(0.1, "cm"),
                                                  gp = grid::gpar(fill = pred.col["FALSE"], col=NA)), 
                                 title = "Without scars", order = 2, 
                                 theme = theme(legend.title = element_text(size = rel(legend.text.size)),
                                               legend.position = "top")),
           No.scars = guide_custom(grid::circleGrob(r = unit(0.1, "cm"), 
                                                  gp = grid::gpar(fill = "black", col=NA)), 
                                 title = "Ignoring scars", order = 1, 
                                 theme = theme(legend.title = element_text(size = rel(legend.text.size)),
                                               legend.position = "top"))) +
    scale_x_continuous(expand = c(0, 0)) +
    theme_classic() + 
    theme(legend.position = "top",
          axis.title = element_text(size = rel(0.7)),
          axis.text = element_text(size = rel(0.5)),
          panel.border = element_rect(colour = "black", fill=NA, linewidth = 0.5),
          axis.line = element_line(colour = 'black', linewidth = 0.1),
          axis.ticks = element_line(colour = 'black', linewidth = 0.1))
    
  return(m.plot)
}

```

```{r}
mov.plot("noscars", healed= FALSE)
mov.plot("sel.healed", healed= FALSE)
mov.plot("mov.healed", healed= TRUE)
mov.plot("mov.sel.healed", healed= TRUE)
```

# Map predictions

First create a distance to shore raster. 

```{r cache=TRUE}
rast.grid <- rast(extent = ext(narwhals.avail.sf), nrows=150, ncols=150, crs = crs(land))
landR <- as.numeric(rasterize(land, rast.grid))
landR[is.na(landR)] <- 0

shore <- distance(landR, vect(land.1))/1000
names(shore) <- "distShore"

shore.w <- shore
shore.w[landR == 1] <- NA
```

Prediction map of SSF using the naive method (sensu Signer et al. 2017). Habitat selection coefficient * covariate. Could look at simulation based method, but it gets at the utilization distribution, and it is very challenging to implement in a environment with lots of small fjords. Given that we want to show the selection, the naive method appears more adequate.


```{r}
ssf.map.noscars <- exp(summary(ssf.noscars)$coef$cond["distShore.end",1] * shore.w)
ssf.map.sel.healed.F <- exp((summary(ssf.sel.healed)$coef$cond["distShore.end",1] * shore.w))
ssf.map.sel.healed.T <- exp((summary(ssf.sel.healed)$coef$cond["distShore.end",1] * shore.w) + 
  (summary(ssf.sel.healed)$coef$cond["distShore.end:Healed.Scars",1] * shore.w))
ssf.map.mov.healed <- exp(summary(ssf.mov.healed)$coef$cond["distShore.end",1] * shore.w)
ssf.map.mov.sel.healed.F <- exp((summary(ssf.mov.sel.healed)$coef$cond["distShore.end",1] * shore.w))
ssf.map.mov.sel.healed.T <- exp((summary(ssf.mov.sel.healed)$coef$cond["distShore.end",1] * shore.w) + 
  (summary(ssf.mov.sel.healed)$coef$cond["distShore.end:Healed.Scars",1] * shore.w))


ssf.maps <- rast(list(noscars = ssf.map.noscars, 
                      sel.healed.F = ssf.map.sel.healed.F,
                      sel.healed.T = ssf.map.sel.healed.T,
                      mov.healed = ssf.map.mov.healed, 
                      mov.sel.healed.F = ssf.map.mov.sel.healed.F,
                      mov.sel.healed.T = ssf.map.mov.sel.healed.T))
```


Each plot separately

```{r}
land.col <- "floralwhite"#'beige'
ssf.min.col <- "#ffffcc"
ssf.mid.col <- "#41b6c4"
ssf.max.col <- "#0c2c84"
# For common color bar
val.range <- range(as.data.frame(ssf.maps))
# function for maps
mapping <- function(model.name, col.box){
  map <- ggplot() +
    geom_spatraster(data = ssf.maps[[model.name]]) +
    scale_fill_gradient2(low=ssf.min.col, mid=ssf.mid.col, high=ssf.max.col,
                      midpoint = mean(val.range), limit = val.range, na.value = NA) +
    geom_sf(data=land, fill=land.col) +
    coord_sf(xlim = c(st_bbox(ssf.maps[[model.name]])[1], st_bbox(ssf.maps[[model.name]])[3]),
           ylim = c(st_bbox(ssf.maps[[model.name]])[2], st_bbox(ssf.maps[[model.name]])[4]), expand = FALSE) +
    theme(panel.border = element_rect(colour = col.box, fill=NA, linewidth = 0.5),
          legend.title = element_text(size = rel(legend.text.size)),
          legend.text = element_text(size = rel(0.5)),
          legend.position = "top",
          axis.text = element_text(size = rel(0.5)),
          axis.ticks = element_line(colour = 'black', linewidth = 0.1)) +
    scale_y_continuous(breaks = seq(70, 80, by =1)) + 
    scale_x_continuous(breaks = seq(-90, -80, by = 5)) +
    guides(fill = guide_colorbar(title = "SSF", order = 4,
                                 theme = theme(legend.title = element_text(size = rel(legend.text.size)),
                                               legend.position = "top"))) +
    theme(plot.margin = unit(plot.mar.l, 'lines'))
  return(map)
}
```

```{r}
mapping("sel.healed.F", pred.col["FALSE"])
```

Create legends
```{r}
legend.plot.map <- as_ggplot(get_legend(mapping("mov.sel.healed.F", "black"))) 
legend.plot.mov <- as_ggplot(get_legend(mov.plot("mov.sel.healed", healed= TRUE))) 
```

## Final figure with both movement and habitat selection

```{r Fig_5_alternative_narwhal_res_, fig.height=6, fig.width=6, dev = 'png'}
ggarrange(
  ggarrange(legend.plot.mov, legend.plot.map, nrow=1, widths = c(0.6, 0.4)),
  ggarrange(
    # Column 1 Model without healed scars 
    ggarrange(mov.plot("noscars", healed= FALSE), 
              mapping("noscars", "black"), 
              nrow=1, legend = "none", widths = c(0.55, 0.45),
              labels = c("A", "B"), label.x = c(0.13, 0.27), label.y = 1.05, 
              font.label = list(size = 8, face="plain")),
    ggarrange(mov.plot("sel.healed", healed= FALSE), 
              mapping("sel.healed.F", pred.col["FALSE"]), 
              mapping("sel.healed.T", pred.col["TRUE"]), 
            nrow=1, legend = "none", widths = c(0.55, 0.225 ,0.225),
            labels = c("C", "D", "E"), 
            label.x = c(0.13, rep(0.175, 2)), label.y = c(1.05, rep(0.92,2)),font.label = list(size = 8, face="plain")),
    ggarrange(mov.plot("mov.healed", healed= TRUE),
              mapping("mov.healed", "black"), nrow=1, legend = "none", widths = c(0.55,0.45),
              labels = c("F", "G"), label.x = c(0.13, 0.27), label.y = 1.05,font.label = list(size = 8, face="plain")),
    ggarrange(mov.plot("mov.sel.healed", healed= TRUE), 
              mapping("mov.sel.healed.F", pred.col["FALSE"]),  
              mapping("mov.sel.healed.T", pred.col["TRUE"]), 
            nrow=1, legend = "none", widths = c(0.55, 0.225 ,0.225),
            labels = c("H", "I", "J"), label.x = c(0.13, rep(0.195, 2)), label.y = c(1.05, rep(0.92,2)),
            font.label = list(size = 8, face="plain")),
    ncol=1),
   ncol = 1 , heights = c(0.1, 1)
)
```

# Utilization Distribution 

We can also look at the UD obtained from the step selection analysis. To estimate the UD, we use eq (10) in Potts and Börger (2022) and follow procedure defined in Appendix E of their paper. Specifically, we use the exact method to derive the UD from the movement kernel, involving eigenvalues and eigenvectors. While we already have the selection function for each model, we also need the corresponding movement distribution function to get the movement kernel as defined in Potts and Börger (2022). Since narwhals are marine mammals, we will only consider locations in the raster that correspond to water cells.

```{r get movement kernel, general parameters}
water_cells <- which(values(landR) == 0) # only selects location that are on water
coords_proj <- xyFromCell(rast.grid, water_cells) # extract their corresponding coordinates in the raster
n <- nrow(coords_proj) # Number of locations to evaluate to get the movement kernel that will be a n x n matrix. row is start point and column is end point of the movement.

# Derive distance to shore covariate
shore_vals <- values(shore.w)[water_cells]  # Covariate that is used in all models: distance to shore
shore_vals_end <- matrix(shore_vals, nrow = n, ncol = n, byrow=TRUE) # each column has the same value: corresponding to the distance to shore at the end point (column = endpoint). 
shore_vals_start <- matrix(shore_vals, nrow = n, ncol = n) # each row as same value: corresponding to the distance to shore at the start point. 

# Derive the matrix of step length: row i and col j corresponds to step length from ith location (start) -> jth location (end).
sf_points <- st_as_sf(as.data.frame(coords_proj), coords = c("x", "y"), crs = chosen.proj)
step_length_matrix <- st_distance(sf_points, sf_points)/1000
step_length_matrix <- matrix(as.numeric(step_length_matrix), nrow=nrow(step_length_matrix))

```

The movement depends on the individuals' starting locations, so we use shore_vals_start as the matrix of covariates for the movement parameters. For each model, we now derive the associated movement parameters.

```{r model parameters}
# For Model 1: ssf.map.noscars
shape1 <- shape.ssf(b1_shape, ssf.noscars,
                    shore_vals_start)
scale1 <- scale.ssf(b2_scale, ssf.noscars,
                    shore_vals_start, healed=FALSE) 


# For Model 2: ssf.map.sel.healed.F and ssf.map.sel.healed.T
shape2 <- shape.ssf(b1_shape, ssf.sel.healed,
                    shore_vals_start)
scale2 <- scale.ssf(b2_scale, ssf.sel.healed,
                    shore_vals_start, healed=FALSE) 

# For Model 3: ssf.map.mov.healed
shape3_T <- shape.ssf(b1_shape, ssf.mov.healed,
                      shore_vals_start)
scale3_T <- scale.ssf(b2_scale, ssf.mov.healed,
                      shore_vals_start, healed=TRUE) 

shape3_F <- shape.ssf(b1_shape, ssf.mov.healed,
                      shore_vals_start)
scale3_F <- scale.ssf(b2_scale, ssf.mov.healed,
                      shore_vals_start, healed=FALSE) 


# For Model 4: ssf.map.movsel.healed
shape4_T <- shape.ssf(b1_shape, ssf.mov.sel.healed,
                      shore_vals_start)
scale4_T <- scale.ssf(b2_scale, ssf.mov.sel.healed,
                      shore_vals_start, healed=TRUE) 

shape4_F <- shape.ssf(b1_shape, ssf.mov.sel.healed,
                     shore_vals_start)
scale4_F <- scale.ssf(b2_scale, ssf.mov.sel.healed,
                     shore_vals_start, healed=FALSE)
```

From it, we derive the matrix of step-length distribution for each model.

```{r step.length}
step_prob_matrix1 <- dgamma(step_length_matrix, 
                            shape = shape1, scale = scale1)
step_prob_matrix2 <- dgamma(step_length_matrix, 
                            shape = shape2, scale = scale2)
step_prob_matrix3_T <- dgamma(step_length_matrix, 
                              shape = shape3_T, scale = scale3_T)
step_prob_matrix3_F <- dgamma(step_length_matrix,
                              shape = shape3_F, scale = scale3_F)
step_prob_matrix4_T <- dgamma(step_length_matrix,
                              shape = shape4_T, scale = scale4_T)
step_prob_matrix4_F <- dgamma(step_length_matrix,
                              shape = shape4_F, scale = scale4_F)
```


We can now derive the movement kernel. Note that we do not incorporate turning angles, effectively ignoring directional correlation in the data. This simplification should not introduce minimal bias because the effect of correlation becomes negligible when inferring long-term behaviour. As Hawkes (2009) observes "As sampling is carried out less frequently or after a large number of steps, CRW increasingly comes to resemble uncorrelated random walk (Okubo 1980). That is relevant in exploring the implications of the correlation identiﬁed in movement behaviour because CRW may generate distance distributions similar to or the same as those of uncorrelated random walk over time-scales relevant to dispersal". In other words, over large scales and long time periods, the habitat use patterns of a correlated random walk are similar to those of an uncorrelated walk. Since the UD reflects the stationary equilibrium of habitat use patterns, neglecting the correlation in movement data when estimating it, should only introduce minimal bias.

Additionally, including turning angles would increase the dimensionality of the transition matrix, making it an n x n x M array, where $M$ refers to the number of discrete bins used to represent the possible angles from $0$ to $2\pi$. This would make matrix inversion computationally unfeasible, so ignoring turning angles is both theoretically reasonable and practically necessary.

```{r movement_kernel:model1_noscar}
movkernel_matrix <- step_prob_matrix1*
  exp(summary(ssf.noscars)$coef$cond["distShore.end",1]*shore_vals_end)
col_sums <- colSums(movkernel_matrix)
movkernel_matrix_norm1 <- t(t(movkernel_matrix)) / col_sums
```

```{r sel_healed}
#  Healed False
movkernel_matrix2_F <- step_prob_matrix2*
  exp((summary(ssf.sel.healed)$coef$cond["distShore.end",1] * shore_vals_end))
col_sums <- colSums(movkernel_matrix2_F)
movkernel_matrix2_F_norm <- t(t(movkernel_matrix2_F) / col_sums)

#  Healed True
movkernel_matrix2_T <- step_prob_matrix2*
  exp((summary(ssf.sel.healed)$coef$cond["distShore.end",1] * shore_vals_end) + 
        (summary(ssf.sel.healed)$coef$cond["distShore.end:Healed.Scars",1] * shore_vals_end))
col_sums <- colSums(movkernel_matrix2_T)
movkernel_matrix2_T_norm <- t(t(movkernel_matrix2_T) / col_sums)

```

```{r mov_healed}
#  Healed False
movkernel_matrix3_F <- step_prob_matrix3_F*
  exp(summary(ssf.mov.healed)$coef$cond["distShore.end",1] * shore_vals_end)
col_sums <- colSums(movkernel_matrix3_F)
movkernel_matrix3_F_norm <-  t(t(movkernel_matrix3_F) / col_sums)

#  Healed True
movkernel_matrix3_T <- step_prob_matrix3_T*
  exp(summary(ssf.mov.healed)$coef$cond["distShore.end",1] * shore_vals_end)
col_sums <- colSums(movkernel_matrix3_T)
movkernel_matrix3_T_norm <- t(t(movkernel_matrix3_T) / col_sums)

```


```{r mov_sel_healed}
#  Healed False
movkernel_matrix4_F <- step_prob_matrix4_F*
  exp(summary(ssf.mov.sel.healed)$coef$cond["distShore.end",1] * shore_vals_end)
col_sums <- colSums(movkernel_matrix4_F)
movkernel_matrix4_F_norm <- t(t(movkernel_matrix4_F) / col_sums)

#  Healed True
movkernel_matrix4_T <- step_prob_matrix4_T*
  exp((summary(ssf.mov.sel.healed)$coef$cond["distShore.end",1] * shore_vals_end) + 
        (summary(ssf.mov.sel.healed)$coef$cond["distShore.end:Healed.Scars",1] * shore_vals_end))
col_sums <- colSums(movkernel_matrix4_T)
movkernel_matrix4_T_norm <-  t(t(movkernel_matrix4_T) / col_sums)
```


```{r eigenvector.fx}
get_eigenraster <- function(movkernel_matrix){
# Compute the eigenvector corresponding to the eigenvalue closest to 1
res <- eigs(movkernel_matrix, k = 1, sigma = 1)

eigenvalue <- res$values
eigenvector <- res$vectors 
eigenvector <- eigenvector / sum(eigenvector)
eigen_raster1 <- rast.grid


values(eigen_raster1) <- NA

# Assign eigenvector values to the water cells
eigen_raster1[water_cells] <- Re(eigenvector)  # removes imaginary parts ??? 

# Optionally normalize again if needed
values(eigen_raster1)[water_cells] <- values(eigen_raster1)[water_cells] / 
                                     sum(values(eigen_raster1)[water_cells], na.rm = TRUE)
return(eigen_raster1)
}
```


```{r getUDs}
eigen_raster1 <- get_eigenraster(movkernel_matrix_norm1)
eigen_raster2_F <- get_eigenraster(movkernel_matrix2_F_norm)
eigen_raster2_T <- get_eigenraster(movkernel_matrix2_T_norm)
eigen_raster3_F <- get_eigenraster(movkernel_matrix3_F_norm)
eigen_raster3_T <- get_eigenraster(movkernel_matrix3_T_norm)
eigen_raster4_F <- get_eigenraster(movkernel_matrix4_F_norm)
eigen_raster4_T <- get_eigenraster(movkernel_matrix4_T_norm)
```

```{r plotUDs}
ud.rasters <- list(
  noscars = eigen_raster1,
  sel.healed.F = eigen_raster2_F,
  sel.healed.T = eigen_raster2_T,
  mov.healed.F = eigen_raster3_F,
  mov.healed.T = eigen_raster3_T,
  sel.mov.healed.T = eigen_raster4_T,
  sel.mov.healed.F = eigen_raster4_F
)
# 
# all_min <- min(sapply(rasters, function(r) minmax(r)[1]))
# all_max <- max(sapply(rasters, function(r) minmax(r)[2]))
# cols <- colorRampPalette(c(ssf.min.col, ssf.mid.col, ssf.max.col))(100)
# 
# # Plot each with the same zlim and col
# par(mfrow = c(3,3))  # adjust grid layout
# plot(eigen_raster1, main = "Stationary Distribution: no scar",
#      col = cols, zlim = c(all_min, all_max))
# plot(eigen_raster2_F, main = "Stationary Distribution: sel.healed.F",
#      col = cols, zlim = c(all_min, all_max))
# plot(eigen_raster2_T, main = "Stationary Distribution: sel.healed.T",
#      col = cols, zlim = c(all_min, all_max))
# plot(eigen_raster3_T, main = "Stationary Distribution: mov.healed.T",
#      col = cols, zlim = c(all_min, all_max))
# plot(eigen_raster3_F, main = "Stationary Distribution: mov.healed.F",
#      col = cols, zlim = c(all_min, all_max))
# plot(eigen_raster4_T, main = "Stationary Distribution: mov.sel.healed.T",
#      col = cols, zlim = c(all_min, all_max))
# plot(eigen_raster4_F, main = "Stationary Distribution: mov.sel.healed.F",
#      col = cols, zlim = c(all_min, all_max))
```


Plot with ggplot to combine

```{r UDforFinal}
val.range.UD <- range(as.data.frame(ud.rasters))
mapping.UD <- function(model.name, col.box, plot.mar.l.UD = c(0.045, 0.2, 0.2, 0.2)){
  map <- ggplot() +
    geom_spatraster(data = ud.rasters[[model.name]]) +
    scale_fill_gradient2(low=ssf.min.col, mid=ssf.mid.col, high=ssf.max.col,
                      midpoint = mean(val.range.UD), limit = val.range.UD, na.value = NA) +
    geom_sf(data=land, fill=land.col) +
    coord_sf(xlim = c(st_bbox(ud.rasters[[model.name]])[1], st_bbox(ud.rasters[[model.name]])[3]),
           ylim = c(st_bbox(ud.rasters[[model.name]])[2], st_bbox(ud.rasters[[model.name]])[4]), expand = FALSE) +
    theme(panel.border = element_rect(colour = col.box, fill=NA, linewidth = 0.5),
          legend.title = element_text(size = rel(legend.text.size)),
          legend.text = element_text(size = rel(0.5)),
          legend.position = "top",
          axis.text = element_text(size = rel(0.5)),
          axis.ticks = element_line(colour = 'black', linewidth = 0.1)) +
    scale_y_continuous(breaks = seq(70, 80, by =1)) + 
    scale_x_continuous(breaks = seq(-90, -80, by = 5)) +
    guides(fill = guide_colorbar(title = "UD", order = 4,
                                 theme = theme(legend.title = element_text(size = rel(legend.text.size)),
                                               legend.position = "top"))) +
    theme(plot.margin = unit(plot.mar.l.UD, 'lines'))
  return(map)
}

mapping.UD("sel.healed.F", pred.col["FALSE"])
```

Create legends
```{r}
legend.plot.map.ud <- as_ggplot(get_legend(mapping.UD("sel.healed.F", "black")))
```

```{r Fig_5_narwhal_res_ud, fig.height=6, fig.width=6, dev = 'png'}
ggarrange(
  ggarrange(legend.plot.mov, legend.plot.map.ud, nrow=1, widths = c(0.6, 0.4)),
  ggarrange(
    # Column 1 Model without healed scars 
    ggarrange(mov.plot("noscars", healed= FALSE), 
              mapping.UD("noscars", "black", plot.mar.l.UD = c(0.3, 0.2, 0.5, 0.2 )), 
              nrow=1, legend = "none", widths = c(0.38, 0.62),
              labels = c("A", "B"), 
              label.x = c(0.19, 0.315), label.y = 1.06, 
              font.label = list(size = 8, face="plain")),
    ggarrange(mov.plot("sel.healed", healed= FALSE), 
              mapping.UD("sel.healed.F", pred.col["FALSE"]), 
              mapping.UD("sel.healed.T", pred.col["TRUE"]), 
            nrow=1, legend = "none", widths =  c(0.38, 0.31,0.31),
            labels = c("C", "D", "E"), 
            label.x = c(0.19, rep(0.15, 2)), label.y = c(1.06, rep(1.06,2)),font.label = list(size = 8, face="plain")),
    ggarrange(mov.plot("mov.healed", healed= TRUE), 
              mapping.UD("mov.healed.F", pred.col["FALSE"]), 
              mapping.UD("mov.healed.T", pred.col["TRUE"]), 
            nrow=1, legend = "none", widths =  c(0.38, 0.31,0.31),
            labels = c("F", "G", "H"), 
            label.x = c(0.19, rep(0.15, 2)), label.y = c(1.06, rep(1.06,2)),font.label = list(size = 8, face="plain")),
    ggarrange(mov.plot("mov.sel.healed", healed= TRUE), 
              mapping.UD("sel.mov.healed.F", pred.col["FALSE"]),  
              mapping.UD("sel.mov.healed.T", pred.col["TRUE"]), 
            nrow=1, legend = "none", widths = c(0.38, 0.31,0.31),
            labels = c("I", "J", "K"), 
            label.x = c(0.19, rep(0.15, 2)), label.y = c(1.06, rep(1.06,2)),
            font.label = list(size = 8, face="plain")),
    ncol=1),
   ncol = 1 , heights = c(0.1, 1)
)
```

Looking at difference between the two UDs.

```{r ud.diff.fx}
mapping.UD.diff <- function(raster.dif,
                            col.box = "black", 
                            plot.mar.l.UD = c(0.2, 0.2, 0.2, 0.2)){
  map <- ggplot() +
    geom_spatraster(data = raster.dif) +
    scale_fill_gradient2(low=ssf.min.col, mid=ssf.mid.col, high=ssf.max.col, na.value = NA) +
    geom_sf(data=land, fill=land.col) +
    coord_sf(xlim = c(st_bbox(raster.dif)[1], st_bbox(raster.dif)[3]),
           ylim = c(st_bbox(raster.dif)[2], st_bbox(raster.dif)[4]), expand = FALSE) +
    theme(panel.border = element_rect(colour = col.box, fill=NA, linewidth = 0.5),
          # legend.title = element_text(size = rel(legend.text.size)),
          # legend.text = element_text(size = rel(0.5)),
          # legend.position = "top",
          axis.text = element_text(size = rel(0.5)),
          axis.ticks = element_line(colour = 'black', linewidth = 0.1)) +
    scale_y_continuous(breaks = seq(70, 80, by =1)) + 
    scale_x_continuous(breaks = seq(-90, -80, by = 5)) +
    guides(fill = guide_colorbar(title = "Diff. in UD", order = 4,
                                 theme = theme(legend.title = element_text(size = rel(legend.text.size)),
                                               legend.position = "top"))) +
    theme(plot.margin = unit(plot.mar.l.UD, 'lines'))
  return(map)
}
```


```{r Fig_SupMat4_narwhal_ud_diff_sel, fig.height=3, fig.width=6, dev = 'png'}
mapping.UD.diff(eigen_raster2_T - eigen_raster2_F)
```

```{r Fig_SupMat5_narwhal_ud_diff_mov, fig.height=3, fig.width=6, dev = 'png'}
mapping.UD.diff(eigen_raster3_T - eigen_raster3_F)
```

```{r Fig_SupMat6_narwhal_ud_diff_sel_mov, fig.height=3, fig.width=6, dev = 'png'}
mapping.UD.diff(eigen_raster4_T - eigen_raster4_F)
```


Check the UDs in terms of their correlation to one another. To see how different they rae.

```{r correlation}
cor(values(eigen_raster2_F), values(eigen_raster2_T), "complete.obs") # selection only
cor(values(eigen_raster3_F), values(eigen_raster3_T), "complete.obs") # movement only
cor(values(eigen_raster4_F), values(eigen_raster4_T), "complete.obs") # both


cor(values(eigen_raster4_F),values(eigen_raster3_F),"complete.obs") # movement only vs both mov and selection
cor(values(eigen_raster4_T),values(eigen_raster3_T),"complete.obs") # movement only vs both mov and selection

```

# Session info

```{r session.info}
sessionInfo()
```

# Litterature cited

Avgar, T., Lele, S. R., Keim, J. L., & Boyce, M. S. (2017). Relative selection strength: Quantifying effect size in habitat‐and step‐selection inference. Ecology and Evolution, 7, 5322-5330. https://doi.org/10.1002/ece3.3122


Hawkes, C. (2009). Linking movement behaviour, dispersal and population processes: is individual variation a key? Journal of Animal Ecology, 78, 894-906. https://doi.org/10.1111/j.1365-2656.2009.01534.x


Muff S., Signer J., & Fieberg J. (2020). Accounting for individual-specific variation in habitat-selection studies: Efficient estimation of mixed-effects models using Bayesian or frequentist computation. J Anim Ecol, 89, 80–92. https://doi.org/10.1111/1365-2656.13087

Okubo, A. (1980). Diffusion and ecological problems: mathematical models. Springer, New York.

Potts, J. R., & Börger, L. (2023). How to scale up from animal movement decisions to spatiotemporal patterns: An approach via step selection. Journal of Animal Ecology, 92, 16-29. https://doi.org/10.1111/1365-2656.13832

Signer, J., Fieberg , & Avgar T. (2017). Estimating utilization distributions from fitted step-selection functions. Ecosphere 8, e01771. https://doi.org/10.1002/ecs2.1771 
